GVN-Hoist: Hoisting Computations from Branches

### 2016-08-11 hoist.tex
formalisatio of partition algorithm

\subsection{Hoisting bottom up with optimistic approach}
\label{subsec:optimistic}
In this approach, the goal is to maximise the total number of hoistings in the
entire function.  This algorithm is very useful when optimizing for code-size.
We collect the GVN of all the instructions in the function and iterate on the
list of instructions having identical GVNs. After that we find the common
dominator dominating all such identical computations and perform legality checks
(Section-\ref{subsec:legality}). Many times it is not possible to hoist all the
instructions to one common dominator, due to legality constraints e.g.,
intersecting side-effects or, profitability constraints e.g., hoisting point too
far. In those cases, this algorithmn would partition the list of identical
instructions into subsets which can be partially hoisted to their respective
common dominators. The partition algorithm is described as follows:

\subsubsection{Partition the list of hoisting candidates to maximize hoisting}
\label{subsec:partition}
In order to hoist a subset of identical instructions, we partition the list of
all candidates in a way to maximise the total number of hoistings.  By sorting
the list of all the candidates in the increasing order of their DFS-in numbers,
we make sure that candidates closer in the list have their common dominator
nearby in cases when there are no fully redundant instructions.

Essentially,
abs(DFSIn(I1) - DFSIn(I2)) < abs(DFSIn(I1) - DFSIn(I3))
=> bbDist(I1, nearest-common-dom(I1, I2)) <= bbDist(I1, nearest-common-dom(I1, I3))

In case of fully redundant computations the nearby elements may not be 'closer'
than the separated ones. TODO: Explain.

This means for candidates closer to each other w.r.t. DFS numbers:
\begin{enumerate}
\item would make fewer checks for legality and profitability.
\item hoisting point will be closer so, the intersection of live-range of the
  instruction with other instructions will be minimal.
\end{enumerate}

>>>>

\subsection{Hoisting bottom up with optimistic approach}
\label{subsec:optimistic}
In this approach, the goal is to maximise the total number of hoistings in the
entire function.  This algorithm is very useful when optimizing for code-size.
We collect the GVN of all the instructions in the function and iterate on the
list of instructions having identical GVNs. After that we find the common
dominator dominating all such identical computations and perform legality checks
(Section-\ref{subsec:legality}). Often times it is not possible to hoist all the
instructions to one common dominator, due to legality constraints e.g.,
intersecting side-effects or, profitability constraints e.g., hoisting point too
far. In those cases, this algorithmn would partition the list of identical
instructions into subsets which can be partially hoisted to their respective
common dominators. The partition algorithm is described as follows:

\subsubsection{Partition the list of hoisting candidates to maximize hoisting}
\label{subsec:partition}
In order to hoist a subset of identical instructions, we partition the list of
all candidates in a way to maximise the total number of hoistings.  By sorting
the list of all the candidates in the increasing order of their depth first
search discovery time stamp \cite{clrs} (DFSIn numbers), we make sure that
candidates closer in the list have their common dominator nearby in cases when
there are no fully redundant instructions. Essentially if,

\begin{verbatim}
BasicBlock B1, B2, B3 // B1 != B2 != B3
DFSIn(B1) = depth first discovery time stamp
DFSIn(B1,B2) = abs(DFSIn(B1) - DFSIn(B2))
DFSIn(B1,B3) = abs(DFSIn(B1) - DFSIn(B3))
BBDist(B1,B2) = BBDist(B1, nearest-common-dom(B1, B2))
BBDist(B1,B3) = BBDist(B1, nearest-common-dom(B1, B3))

Then,
DFSIn(B1,B2) < DFSIn(B1,B3) implies
 BBDist(B1,B2) <= BBDist(B1,B3)
\end{verbatim}

Same property would hold for instructions I1, I2, I3 provided they are not fully
redundant i.e., none of the basic blocks containing I1, I2 and I3 dominate the
other.

\begin{verbatim}
DFSIn(I1) = DFSIn of containing basic block
DFSIn(I1,I2) = abs(DFSIn(I1) - DFSIn(I2))
DFSIn(I1,I3) = abs(DFSIn(I1) - DFSIn(I3))
BBDist(I1,I2) = BBDist(I1, nearest-common-dom(I1, I2))
BBDist(I1,I3) = BBDist(I1, nearest-common-dom(I1, I3))

Then,
 DFSIn(I1,I2) < DFSIn(I1,I3) implies
  BBDist(I1,I2) <= BBDist(I1,I3)
\end{verbatim}

It says that if I1 is closer to I2 than I3 in the list of candidates sorted by
DFSIn numbers, then the nearest common dominator of I1 and I2 will be closer (in
terms of basic block distance) to I1, than nearest common dominator of I1 and
I3.

In the presence of fully redundant computations in the list of hoistable
candidates this equation may not hold. For example, if there is another
instruction I4 which is dominated by I1, but I4 is present in a basic block
which is farther than DFSIn(I1, I2) i.e., $DFSIn(I1,I2) < DFSIn(I1,I4)$;
however, $BBDist(I1,I4) = 0 < BBDist(I1,I2) >=1$

So by sorting candidates w.r.t. their DFSIn numbers:
\begin{enumerate}
\item would make fewer checks for legality and profitability.
\item hoisting point will be closer so, the intersection of live-range of the
  instruction with other instructions will be minimal.
\end{enumerate}
